When does vapor pressure deficit drive or reduce evapotranspiration?

### 2018-03-26 doc/paper/vpd_et_paper.tex
done with discussion, need to add convex vs convave discussion

\begin{figure}
\centering
\centerline{\includegraphics[width=1.4\textwidth]{./swc_boxplot.pdf}}
\caption{$\sigma \cdot uWUE$ as a function of PFT and SWC. Each box plot represents 5\% of all observations.}
\label{swc_boxplot}
\end{figure}

@@

To summarize, our theory is limited by its inability to account for soil water impacts on land surface response, and feedbacks between SWC and VPD. Fortunately, for most PFTs SWC's effect on ET is relatively rare (<30\% of observations) and does not manifest itself in the majority of observations and bulk statistics. However, for GRA SWC has a strong effect decreasing water use efficiency for the majority of the observations. Soil moisture effects explain the deficiencies of our theory in Section \ref{testing}, particularly for GRA. By conceptualizing SWC effects as a change in $uWUE$ (or $g_1$ as well), it will be possible for future analysis to explore the importance of soil moisture on plant response to VPD.

@@

Our theory suggests that for a majority of environmental conditions, plants will tend to conserve water and reduce ET with increasing VPD. Stomatal regulation and plant physiological response strongly regulate ET, and this regulation varies by PFT. CROs are the least water conservative, while DBF, EBF, ENF, GRA and CSH are progressively more water conservative (more likely to reduce ET in response to increasing VPD). Observations of ET response to VPD exhibit the same general behavior as the theory, with ET response becoming more positive (atmospheric demand dominating) as environmental VPD increases within a PFT, and more negative for PFTs that are adapted to arid conditions and prioritize water conservation over primary production.

@@

Lastly, our paper builds important intuition for how plants will respond to global change-induced shifts in VPD, which should increase in the future. Plant physiological responses to direct CO$_2$ effects \citep[e.g.,][]{Swann_2016} receives more attention than physiological response to indirect effects like increased VPD. Here, we provide broad PFT-focused results showing a likely decrease in ET in response to positive VPD perturbations (atmospheric drying). Additionally, we derive an explicit conceptual framework for ET response as a function of two plant type parameters, $g_1$ and uWUE (e.g. Equation \ref{et}). This framework can be used to understand ET response to a range of environmental change, and forms a foundation for resolving varying response to more nuanced consideration of plant type and climate. Any plant physiological heterogeneity that can be conceptualized with shifts in $g_1$ \citep[e.g.][]{Lin_2015} and/or uWUE \citep[e.g.][]{Zhou_2014} is representable within our framework, which opens the door for a hierarchy of more sophisticated climate- and plant-specific analysis of ET sensitivity to environmental variables (including VPD). We argue that such simplified conceptual frameworks are critical tools for disentangling land-atmosphere feedbacks at various scales, from diurnal to seasonal and beyond, and to characterize ET response in a warmer, atmospherically drier, and enriched CO$_2$ world.

>>>>

\begin{figure}
\centering
\centerline{\includegraphics[width=1.4\textwidth]{./swc_boxplot.pdf}}
\caption{$\sigma \cdot uWUE$ as a function of PFT and SWC. Each box plot represents 5\% of all observations. Note that the highest 10\% of SWC observations were excluded from the WSA and SAV figures, to better resolve high variability in narrow low-SWC bins for those PFTs.}
\label{swc_boxplot}
\end{figure}

@@

To summarize, our theory is limited by its inability to account for soil water impacts on land surface response, and feedbacks between SWC and VPD. Fortunately, for most PFTs SWC's effect on ET is relatively rare (<30\% of observations) and does not manifest itself in the majority of observations and bulk statistics. However, for GRA SWC has a strong effect decreasing water use efficiency for the majority of the observations. Soil moisture effects explain the deficiencies of our theory in Section \ref{testing}, particularly for GRA. By conceptualizing SWC effects as a change in $uWUE$ (and/or $g_1$), it will be possible for future analysis to explore the importance of soil moisture on plant response to VPD.

@@

Our theory suggests that for a majority of environmental conditions, plants will tend to conserve water and reduce ET with increasing VPD. Stomatal regulation and plant physiological response strongly regulate ET, and this regulation varies by PFT. CROs are the least water conservative, while DBF, EBF, SAV, GRA, MF, WSA, ENF and CSH are progressively more water conservative (more likely to reduce ET in response to increasing VPD). SAV and WSA exhibit positive ET response to VPD not necessarily because of poor water conservation stragegies relative to other PFTs, but because of greater occurence of high atmospheric demand (VPD) relative to other PFTs. Observations of ET response to VPD exhibit the same general behavior as the theory, with ET response becoming more positive (atmospheric demand dominating) as environmental VPD increases within a PFT, and more negative for PFTs that are adapted to arid conditions and prioritize water conservation over primary production.

@@

Lastly, our paper builds important intuition for how plants will respond to VPD pertubations, including expected increases in VPD induced by global change.  Plant physiological responses to direct CO$_2$ effects \citep[e.g.,][]{Swann_2016} receives more attention than physiological response to indirect effects like increased VPD. Here, we provide broad PFT-focused results showing a likely decrease in ET in response to positive VPD perturbations (atmospheric drying), which is consistent with recent observational analysis \citep[e.g.,][]{Rigden_2017}. Feedbacks between the land and the atmosphere may alter the net response to a global VPD perturabation, but our focus on the one way plant response to a VPD perturbation in the atmopsheric boundary layer is an important first step to disentangling such feedbacks, both in observations and model simulations of the present and future. By removing Penman Monteith's dependence on implicit relationships between GPP, VPD, and ET, we allow for explicit, accuruate analysis of plant-VPD feedbacks in the atmospheric boundary layer (Equation \ref{et}). Our approach can be extended to examine varying plant response to more nuanced consideration of plant type and climate. Any plant physiological heterogeneity that can be conceptualized with shifts in $g_1$ \citep[e.g.][]{Lin_2015, Medlyn_2017} and/or uWUE \citep[e.g.][]{Zhou_2014} is representable within our framework, which opens the door for a hierarchy of more sophisticated climate- and plant-specific analysis of ET sensitivity to environmental variables (including VPD). We argue that such simplified conceptual frameworks are critical tools for disentangling land-atmosphere feedbacks at various scales, from diurnal to seasonal and beyond, and to characterize ET response in a warmer, atmospherically drier, and enriched CO$_2$ world.
